---
title: "SQL and R"
author: "Amanda Arce"
date: "September 8, 2018"
output: html_document
---


###Choose six recent popular movies. Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5. Take the results (observations) and store them in a SQL database. Load the information into an R dataframe.


###Below I will connect MySQL database "popular" to R Studio


```{r}

library(RMySQL)
library(dbConnect)
library(ggplot2)
library(dplyr)




mydb = dbConnect(MySQL(), user='root', password='haley', dbname='popular', host='localhost')


popular <- dbGetQuery(mydb, "select * from reviewers")

popular

```
###Above I have queried the reviewers table, followed by the ratings table below.
```{r}

popular <- dbGetQuery(mydb, "select * from ratings")

popular
```
###Here I have queried the popular movies table

```{r}
popular <- dbGetQuery(mydb, "select * from popular_movies ")

popular

```


###At this point, I will now create a join query to analyze the title of the movies and how rated them.
```{r}


popular <- dbGetQuery(mydb,"SELECT user_id, title, ratings.movie_id, rating
                    FROM popular_movies
                    JOIN ratings
                     on popular_movies.movie_id = ratings.movie_id")

popular
```


###On this next query, I decided to add the "Reviwers" table to bring in the "Raters names"
```{r}
popular <- dbGetQuery(mydb,
                    "
SELECT 
    ratings.user_id,raters_name, title, rating
FROM
    ratings
JOIN
    popular_movies ON popular_movies.movie_id = ratings.movie_id
JOIN
    reviewers ON reviewers.user_id = ratings.user_id
ORDER BY 
    rating DESC
")

popular
```

### Now I will see what the average movie rating were from my raters.
```{r}
ggplot(data = popular, aes(popular$title, popular$rating)) + 
  stat_summary(fun.y = mean,
               geom = "bar", aes(fill = popular$title)) +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  theme(legend.position="none") +
  xlab("Movie Name") +
  ylab("Rating") +
  ggtitle("Avg Rating for Movies")
```

###Conclusion: 
###Based on my analysis, my friends have chosen Black Panther to be one of the highest ratings out of the six movies; followed by The Avengers.  