---
title: "Final project"
author: "Amanda Arce"
date: "December 4, 2018"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

#Libraries


```{r message=FALSE, warning=FALSE, paged.print=FALSE}

library(rmdformats)
library(readr)
library(tidyverse)
library(dplyr)
library(psych)
library(knitr)
library(ggthemes)
library(kableExtra)
library(plotly)
library(forecast)
```

#Introduction 

Mental illness is a serious disease that affects millions of people worldwide.  One unfortunate outcome is some individuals decide to commit suicide.  This study will attempt to understand suicide rates for different countries, and sex.

#Data

The data is taken from kaggle.com's https://www.kaggle.com/szamil/who-suicide-statistics/home.  The dataset consists of the following variables: country, year, sex, age, suicides_no, and population. It also consists of 43776 rows of data. This dataset was aggregated from multiple datasets to include only the previous variables.

```{r}
suicide <- read_csv("https://raw.githubusercontent.com/mandiemannz/Data-607--Fall-18/master/who_suicide_statistics.csv")
head(suicide)
```


#Exploratory Data Analysis



A preiminilary analysis shows that the # of people in each age group are distributed evenly: 
```{r}
agecount <- suicide %>%
  group_by(age) %>%
  summarize(n=n())

agecount
```

```{r}
ggplot(agecount, aes(agecount$age, agecount$n)) +
   stat_summary(fun.y = sum,
               geom = "bar") 
```


The # of values for each country in the data varies.  The min of data recorded appears 12 times, while the max for a country appears 456 times.  Some countries have more recorded data than others-  this can be due to the size of the country, its population, the ease of acquiring that data.
```{r}
countrycount <- suicide %>%
  group_by(country) %>%
  summarize(n=n())

summary(countrycount)
```



### If we filter the dataset by number of suicides greater than 1000, the data becomes easier to see which countries have the greatest sucidie numbers 

```{r}
countrysuicide <- suicide %>%
  filter(suicides_no > 1000) 

p <- ggplot(data = countrysuicide, aes(reorder(countrysuicide$country, -countrysuicide$suicides_no), countrysuicide$suicides_no)) + 
  stat_summary(fun.y = sum,
               geom = "bar", fill = "grey") +
  theme_fivethirtyeight() +
  xlab("Country") + 
  ylab("Count") +
  coord_flip() 

ggplotly(p)
```
It appears that the Russian Federation has the most # of suicides (1450349), followed by the United States of America (1138176)


### We can also break down by gender and country
```{r}
countrysuicide <- suicide %>%
  filter(suicide$suicides_no > 1000)

p <- ggplot(data = countrysuicide, aes(countrysuicide$`sex`,countrysuicide$suicides_no)) + 
  stat_summary(fun.y = sum,
               geom = "bar", aes(fill=sex)) +
  xlab("Sex") + 
  ylab("Count") +
  facet_wrap(~countrysuicide$country) +
  ggtitle("Suicide vs. Country and Sex")

ggplotly(p)
```
According to this graph, in both Russa and the United States, Males outnumber the women in suicide #s.



##Age Group

###United States

```{r}
countrysuicide <- suicide %>%
  filter(country == "United States of America")

p <- ggplot(data = countrysuicide, aes(countrysuicide$age,countrysuicide$suicides_no)) + 
  stat_summary(fun.y = sum,
               geom = "bar", aes(fill=sex)) +
  xlab("Age") + 
  ylab("Count") +
  ggtitle("Suicide vs. Country and Sex") 

ggplotly(p)
```

###Russian Federation
```{r}
countrysuicide <- suicide %>%
  filter(country == "Russian Federation")

p <- ggplot(data = countrysuicide, aes(countrysuicide$age,countrysuicide$suicides_no)) + 
  stat_summary(fun.y = sum,
               geom = "bar", aes(fill=sex)) +
  xlab("Age") + 
  ylab("Count") +
  ggtitle("Suicide vs. Country and Sex")

ggplotly(p)
```

In both the United States and Russia, Males outnumber the females in sucidie.  The age group most vulnerable is the 35-54 age group.



##Suicide over the years in the Dataset

```{r}
p <- ggplot(data = suicide, aes(suicide$year, suicide$suicides_no)) + 
  stat_summary(fun.y = sum,
               geom = "bar", aes(fill=year)) +
  theme(legend.position="none") +
  xlab("Year") +
  ylab("Count") +
  ggtitle("Suicide") 

ggplotly(p)

```

The data shows an increasing trend until around 2000, where it beings to flatten and slow decrease.  The year 2016 has a large difference - it seems as if the data for 2016 is incomplete.

```{r}
tssuicide <- suicide %>%
  dplyr::select(year, suicides_no) 

tssuicide$suicides_no[is.na(tssuicide$suicides_no)] <- 0

tssuicide <- tssuicide %>% 
  group_by(year) %>%
  summarize_all(sum) 

tssuicide$year <- as.character(tssuicide$year)

myts <- ts(tssuicide, start = 1985, end = 2016, frequency = 1)

autoplot(myts) 
```



## Population #s
```{r}
p <- ggplot(data = suicide, aes(suicide$year, suicide$population)) + 
  stat_summary(fun.y = sum,
               geom = "bar", aes(fill=year)) +
  theme(legend.position="none") +
  xlab("Year") +
  ylab("Count") +
  ggtitle("Suicide") 

ggplotly(p)
```

Throughout the dataset, you can see a steady increasing trend in population

#ANOVA Model

We can run an ANOVA test on different data within our dataset to see if there is a statisticaly significant difference between the years, and countries in the data.

$$H_0: \mu_{1985} = \mu_{1986} = \mu_1{987} ... = \mu_{2016}$$

$$H_a: \mu_{1985} \neq \mu_{1986} \neq \mu_{1987} ... \neq \mu_{2016} $$


```{r}
model <- lm(suicides_no ~ year, data= suicide)

anova(model)
```


Since the p-value is low (less than 0.05) - null must go; meaning, the $H_0$ is rejected and the $H_a$ is accepted.  There is significant evidence that number of suicides is different throughout the years





$$H_0: \mu_{country1} = \mu_{country2} = \mu_1{country3} ... = \mu_{countryn}$$



$$H_a: \mu_{country1} \neq \mu_{country2} \neq \mu_{country3} ... \neq \mu_{countryn} $$


```{r}
model <- lm(suicides_no ~ country, data= suicide)

anova(model)
```

#Conclusions 

  After analyzing the suicide data from Kaggle, the data shows that on average, Males are more likly to commit suicide than females, the number of suicides  over the years has increased, the number of population overall has also increased. 

  Russia and the United States have the highest sucide rates within the dataset.  In both the United States and Russia, Males outnumber the females in  suicide The age group most vulnerable is the 35-54 age group.


Reference:

https://www.kaggle.com/szamil/who-suicide-statistics















