---
title: "CUNY MSDS DATA 607 HW 3"
author: "Amanda Arce"
date: "September 16, 2018"
output: html_document
---
```{r}

library(stringr)
library(kableExtra)
library(knitr)

```
3.) Copy the introductory example. The vector name stores the extracted names.

a. Use the tools of this chapter to rearrange the vector so that all elements conform to the        standard first_name last_name.
b. Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).

c. Construct a logical vector indicating whether a character has a second name.

```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
 
```
a. Use the tools of this chapter to rearrange the vector so that all elements conform to the        standard first_name last_name.
```{r}
#extract words
originalNames <-  unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))

originalNames
```
```{r}

#extract middle names 
names1 <- str_replace(originalNames, "\\s[A-z]\\. ", " ")
names1
```
```{r}
# change first name to last- found word groups then changed them 
names2 <- str_replace(names1, "(\\w+),\\s(\\w+)", "\\2 \\1")
names2

```



```{r}

#Removed title of name of Rev. Timothy Lovejoy, Dr. Julius Hibbert
newNames <- str_replace(names2, "[A-z]{2,3}\\. "," ")

newNames
```


```{r}
originalN <- data.frame(originalNames)
newN <- data.frame(newNames)

```

```{r}
# create a table with Kable to list the differences 
kable(list(originalN,newN),caption='Names')

```
(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).
```{r}
title <- str_detect(names2,"[A-z]{2,3}\\. ")
df1 <- data.frame(names2, title)

df1

```
(c) Construct a logical vector indicating whether a character has a second name.

```{r}
title <- str_detect(originalNames,"[A-Z]{1}\\." )
df2 <- data.frame(originalNames, title)

df2
```

4.) Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

a   [0-9]+\$
b. \b[a-z]{1,4}\b
c. *?\.txt$
d. \d{2}/\d{2}/\d{4}
e. <(.+?)>.+?</\1>  

a.  [0-9]+\$ : Pick numbers 0-9,+ - one or more time, $
```{r}
ex_one <- "15689142$now!@23$"
str_extract_all(ex_one, "[0-9]+\\$" )



```
b. \b[a-z]{1,4}\ //b Word Edge, [a-z] letters a-z are lowercase, {1,4} min,max chars 
```{r}
ex_two <- " Today is a day great"
str_extract_all(ex_two, "\\b[a-z]{1,4}\\b")
```
c. .*?\.txt$  takes out files: . any chars except line break, /* zero or more times, //. literal ., $ end of line

```{r}


ex_three <- "c:/local/amanda/homework.txt"
str_extract_all(ex_three, ".*?\\.txt$")
```
d. \d{2}/\d{2}/\d{4}  takes out dates: //d any num, {2} 2 characters, / a '/'
```{r}
ex_four <-  "07/02/1983"

str_extract_all(ex_four, "\\d{2}/\\d{2}/\\d{4}")
  
```
e. <(.+?)>.+?</\1> takes out html code:

```{r}
ex_five <- "<p>What kind of line is this, a paragraph</p>"
str_extract_all(ex_five, "<(.+?)>.+?</\\1>")

```

9. The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others!

```{r}
data <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
#takes out all capitalperiods and letters 
mystery <- unlist(str_extract_all(data, "[[A-Z].]"))
mystery
```

```{r}
# Here I joined the letters together, and removed the spaces
mystery <- paste(mystery, collapse = "") 

#replaced the '.'s with a space
str_replace_all(mystery, "[.]", " ")
```