---
title: "CUNY MSDS DATA 607 Assignment 5"
author: "Amanda Arce"
date: "September 29, 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: monochrome


---

```{r}

library(tidyr)
library(dplyr)
library(tidyverse)
library(knitr)

```
```{r}

airlines <- read.csv("https://raw.githubusercontent.com/mandiemannz/Data-607--Fall-18/master/flights.csv", stringsAsFactors = FALSE)

airlines

```

Data wrangling added new columns: total_flights, percent_ontime, and percent_delayed
```{r}



airline_dt <-airlines
airline_df <-  airline_dt %>%
  gather("loc","flights ", 3:7) %>%
  spread(Status, "flights ") %>%
  mutate(total_flights = delayed + on_time) %>%
  mutate(percent_ontime = on_time / total_flights ) %>%
  mutate(percent_delayed = delayed / total_flights) 

airline_df
```
```{r}
airline_df %>%
  group_by(Carrier)%>%
  summarize(total_flights = sum(total_flights))
```


Now let us see which airline and city has the highest percentage of on-time flights

```{r}
airline_df %>%
  filter(percent_ontime== max(percent_ontime))
```
Alaska airline with flights from Phoenix had the highest on-time percentage from looking at the entire dataset. Percent of 0.95

Now we are going to check which airline and city has the worst percentage of on-time flights


```{r}
airline_df %>%
  filter(percent_delayed== max(percent_delayed))


```
AM West airlines with flights from San Francisco had the most delayed percentage from looking at the entire dataset. Percent of 0.28 and are on-time 0.71 percent. 

Visualization
```{r}
ggplot(airline_df, aes(airline_df$on_time, airline_df$delayed)) +
  geom_point(aes(color = airline_df$Carrier)) + 
  ggtitle("Num of flights on-time vs. delayed for both Airlines") +
  xlab("On-time flights") +
  ylab("Delayed flights") +
  theme_dark()
```